### 2018-02-01 submission/manuscript.Rmd
result section edits

***Number of OTUs is Dependent on the HiFi DNA Polymerase Used:***
In order to compare the number of OTUs across individuals a Z-score normalization by individual, by cycle number, was employed on the number of OTUs data. After normalization, we identified that there was a HiFi DNA polymerase dependent difference that was consistent across sub-sampling [Figure 1]. Lower cycle numbers (15-20) tended to result in less differences between HiFI DNA polymerase while cycle numbers of 25, 30, and 35 tended to have more distinct differences [Figure 1]. All sub-sampling levels were significantly different only at 35 cycles (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "35x")[, "bh"]) < 0.0001, invisible("<"), invisible("check data"))` 0.0001) [Table S1]. There were differences in HiFi DNA polymerase at 25 and 30 cycles but the sub-sampling depth had to be 5000 or higher (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "30x", sub_sample_level != 1000)[, "bh"]) < 0.05 & max(filter(fs_anova_summary, cycle == "25x", sub_sample_level != 1000)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S1]. Platinum HiFi DNA polymerase was the main driver of the differences observed across all sub-sampling depths at 35 cycles, based on a Tukey post-hoc test (P-value `r ifelse(max(filter(fs_tukey_summary, padj < 0.05)[, "padj"]) < 0.05,  invisible("<"), invisible("check data"))` 0.05) [Table S2].

@@

***Minimal Bray-Curtis Differences are Detected by Cycle Number:***
Overall, there was very little difference between each respective 5-cycle increment (e.g. 15x vs 20x) for both fecal and mock samples and this was consistent across the different sub-samplings used [Figure 3]. There were only two differences between 5-cycle increments that were identified. First, there were differences for the same fecal sample between 20x vs. 25x that was independent of HiFi DNA polymerase but dependent on sub-sampling depth (sub-sampled to 1000 = `r round(median(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`) (median (25% - 75% quantile)), sub-sampled to 5000 = `r round(median(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`), sub-sampled to 10000 = `r round(median(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`)) [Figure 3A]. Second, for the mocks, where data is available, there were larger difference between 20x and 25x (sub-sampled to 1000 = `r round(median(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`)) [Figure 3B]. However, these differences between the next 5 cycle increment do not persist once 25 cycles are reached [Figure 3].

Using PERMANOVA to test for differences within HiFi DNA polymerase groups based on cycle number, only Phusion had cycle dependent differences at 1000 and 5000 sub-sampling depth (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)` and `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 5000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`, respectively). Phusion was one of only two HiFi DNA polymerases that managed to have samples for the 1000 sub-sampling depth at 15 cycles. Next, we assessed whether there were any major differences between 5 cycle increments within each sample. We found that there was no detectable difference in Bray-Curtis index when comparing to the previous 5 cycle increment (P-value `r ifelse(min(bray_sim_to_prev_cycle_kruskal_results$bh, na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05). However, Phusion at 1000 sub-sampling depth had a P-value = `r format(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000, taq == "PHU")[, "pvalue"], digits = 1, nsmall = 1)` before multiple comparison correction. It should be noted that at higher sub-sampling depths these differences in Bray-Curtis indices disappear for both differences in cycle number and within 5 cycle increments within an individual.

***Sequence Error is Dependent on both HiFi DNA Polymerase and Cycle Number:***
Differences by HiFi DNA polymerase in the median average per base error varied without a clear pattern across sub-sampling depth [Table S5]. Generally, the highest per base median average error rates were for the Kappa HiFI DNA polymerase [Figure 4]. This error rate was minimally affected by both the pre-cluster step and chimera removal by VSEARCH [Figure 4]. There were small differences in the per base error rate between the various HiFi DNA polymerase at lower cycle numbers and larger differences at higher cycle number with Platinum having the largest differences of all the HiFi DNA polymerase [Figure 4B-C and Table S6].

The total sequences with at least one error had multiple differences at different cycle numbers and was mostly alleviated by the use of the pre.cluster step [Figure S1]. Major differences before this pre.cluster step were driven by large differences in Accuprime^TM^ and Platinum versus the other HiFi DNA polymerase tested [Figure S1 & Table S7 & S8]. Although Accuprime^TM^ had the lowest per base error rate it had the largest number of sequences with at least one error, regardless of cycle number or sub-sampling depth [Figure S1]. However, this increased number of sequences with an error can be drastically lowered with existing bioinformatic approaches [Figure S1]. Investigation of whether there were HiFi DNA polymerase dependent effects on base substitution found that there were generally no biases in the types of substitution made [Figure S2].

***Chimeric Sequences Correlate with OTUs and are HiFi DNA Polymerase Dependent:***
After chimera removal using VSEARCH and removal of sequences that did not classify as bacteria we assessed the percentage of sequences that were still chimeric within our mock community. At all levels of sub-sampling and cycle number there were significant differences between the HiFi DNA polymerase used (P-value `r ifelse(max(filter(kruskal_chimera_prev, sub_sample_level > 500)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S9]. Differences between Platinum and all other HiFi DNA polymerases accounted for the vast majority of these differences independent of cycle number and sub-sampling depth when using a Dunn's post-hoc test [Table S10]. Generally, across sub-sampling depth and cycle number Accuprime^TM^ had the lowest chimera prevalence of all the HiFi DNA Polymerases regardless of whether pre.cluster or VSEARCH had been used [Figure 5].

For all DNA polymerase, a positive correlation was observed between chimeric sequences and number of OTUs, with this correlation being strongest for Accuprime, Platinum and Phusion HiFi DNA Polymerase [Figure 6]. In general, the R^2^ value between the number of OTUs and chimeric sequences did not change from the use of pre.cluster and VSEARCH [Figure 6]. Taken together, this data suggests that a strong correlation exists between the number of OTUs and the prevalence of chimeric sequences.

>>>>

***The Number of OTUs is Dependent on the HiFi DNA Polymerase Used:***
In order to compare the number of OTUs across individuals a Z-score normalization by individual, by cycle number, was employed on the number of OTUs data. After normalization, we identified that there was a HiFi DNA polymerase dependent difference that was consistent across sub-sampling [Figure 1]. Lower cycle numbers (15-20) tended to result in less differences between HiFI DNA polymerase while cycle numbers of 25, 30, and 35 tended to have more distinct differences [Figure 1]. All sub-sampling levels were significantly different only at 35 cycles (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "35x")[, "bh"]) < 0.0001, invisible("<"), invisible("check data"))` 0.0001) [Table S1]. There were differences in HiFi DNA polymerase at 25 and 30 cycles but the sub-sampling depth had to be 5000 or higher (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "30x", sub_sample_level != 1000)[, "bh"]) < 0.05 & max(filter(fs_anova_summary, cycle == "25x", sub_sample_level != 1000)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S1]. Platinum HiFi DNA polymerase was the main driver of the differences observed across all sub-sampling depths at 35 cycles, based on a Tukey post-hoc test (P-value `r ifelse(max(filter(fs_tukey_summary, padj < 0.05)[, "padj"]) < 0.05,  invisible("<"), invisible("check data"))` 0.05) [Table S2].

@@

***Minimal Bray-Curtis Differences are Detected by Cycle Number:***
Overall, there was little difference between each respective 5-cycle increment at higher cycle numbers, for both fecal and mock samples, and this was consistent across the different sub-samplings [Figure 3]. There were only two differences between 5-cycle increments that were identified. First, there were differences for the same fecal sample between 20 versus 25 cycles that was independent of HiFi DNA polymerase but dependent on sub-sampling depth (sub-sampled to 1000 = `r round(median(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`) (median (25% - 75% quantile)), sub-sampled to 5000 = `r round(median(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`), sub-sampled to 10000 = `r round(median(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`)) [Figure 3A]. Second, for the mocks, where data is available, there were larger difference between 20 and 25 cycles (sub-sampled to 1000 = `r round(median(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`)) [Figure 3B]. However, these differences between the next 5-cycle increment do not persist once 25 cycles are reached [Figure 3].

Using PERMANOVA to test for differences within HiFi DNA polymerase groups based on cycle number, only Phusion had cycle dependent differences at 1000 and 5000 sub-sampling depth (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)` and `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 5000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`, respectively). Phusion was one of only two HiFi DNA polymerases that managed to have samples for the 1000 sub-sampling depth at 15 cycles. Next, we assessed whether there were any major differences between 5-cycle increments within each sample. We found that there was no detectable difference in Bray-Curtis index when comparing to the previous 5-cycle increment (P-value `r ifelse(min(bray_sim_to_prev_cycle_kruskal_results$bh, na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05). However, Phusion at 1000 sub-sampling depth had a P-value = `r format(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000, taq == "PHU")[, "pvalue"], digits = 1, nsmall = 1)` before multiple comparison correction. It should be noted that at higher sub-sampling depths these differences in Bray-Curtis indices disappear for both differences in cycle number and within 5-cycle increments within an individual.

***Sequence Error is Dependent on both HiFi DNA Polymerase and Cycle Number:***
Differences by HiFi DNA polymerase in the median average per base error varied without a clear pattern across sub-sampling depth [Table S5]. Generally, the highest per base median average error rates were for the Kappa HiFi DNA polymerase [Figure 4]. This error rate was minimally affected by both the 'pre.cluster' step and chimera removal by VSEARCH [Figure 4]. There were small differences in the per base error rate between the various HiFi DNA polymerase at lower cycle numbers and larger differences at higher cycle number with Platinum having the largest differences of all the HiFi DNA polymerase [Figure 4B-C and Table S6].

The total sequences with at least one error had multiple differences at different cycle numbers and was mostly alleviated by the use of the 'pre.cluster' step [Figure S1]. Major differences before this 'pre.cluster' step were driven by large differences in Accuprime^TM^ and Platinum versus the other HiFi DNA polymerase tested [Figure S1 & Table S7 & S8]. Although Accuprime^TM^ had the lowest per base error rate it had the largest number of sequences with at least one error, regardless of cycle number or sub-sampling depth [Figure S1]. However, this increased number of sequences with an error can be drastically lowered with existing bioinformatic approaches [Figure S1]. Investigation of whether there were HiFi DNA polymerase dependent effects on base substitution found that there were generally no biases in the types of substitution made [Figure S2].

***Chimeric Sequences are HiFi DNA Polymerase Dependent and Correlate with Number of OTUs:***
After chimera removal using VSEARCH and removal of sequences that did not classify as bacteria, we assessed the percentage of sequences that were still chimeric within our mock community. At all levels of sub-sampling and cycle number there were significant differences between the HiFi DNA polymerase used (P-value `r ifelse(max(filter(kruskal_chimera_prev, sub_sample_level > 500)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S9]. Differences between Platinum and all other HiFi DNA polymerases accounted for the vast majority of these differences independent of cycle number and sub-sampling depth when using a Dunn's post-hoc test [Table S10]. Generally, across sub-sampling depth and cycle number Accuprime^TM^ had the lowest chimera prevalence of all the HiFi DNA Polymerases regardless of whether 'pre.cluster' or chimera removal with VSEARCH had been used [Figure 5].

For all DNA polymerase, a positive correlation was observed between chimeric sequences and number of OTUs, with this correlation being strongest for Accuprime, Platinum and Phusion HiFi DNA Polymerase [Figure 6]. The R^2^ value between the number of OTUs and chimeric sequences did not change from the use of 'pre.cluster' and the removal of chimeras with VSEARCH [Figure 6]. Taken together, this data suggests that a strong correlation exists between the number of OTUs and the prevalence of chimeric sequences.

### 2018-02-01 submission/manuscript.tex
result section edits

\textbf{\emph{Number of OTUs is Dependent on the HiFi DNA Polymerase
Used:}} In order to compare the number of OTUs across individuals a
Z-score normalization by individual, by cycle number, was employed on
the number of OTUs data. After normalization, we identified that there
was a HiFi DNA polymerase dependent difference that was consistent
across sub-sampling {[}Figure 1{]}. Lower cycle numbers (15-20) tended
to result in less differences between HiFI DNA polymerase while cycle
numbers of 25, 30, and 35 tended to have more distinct differences
{[}Figure 1{]}. All sub-sampling levels were significantly different
only at 35 cycles (P-value \textless{} 0.0001) {[}Table S1{]}. There
were differences in HiFi DNA polymerase at 25 and 30 cycles but the
sub-sampling depth had to be 5000 or higher (P-value \textless{} 0.05)
{[}Table S1{]}. Platinum HiFi DNA polymerase was the main driver of the
differences observed across all sub-sampling depths at 35 cycles, based
on a Tukey post-hoc test (P-value \textless{} 0.05) {[}Table S2{]}.

@@

\textbf{\emph{Minimal Bray-Curtis Differences are Detected by Cycle
Number:}} Overall, there was very little difference between each
respective 5-cycle increment (e.g.~15x vs 20x) for both fecal and mock
samples and this was consistent across the different sub-samplings used
{[}Figure 3{]}. There were only two differences between 5-cycle
increments that were identified. First, there were differences for the
same fecal sample between 20x vs.~25x that was independent of HiFi DNA
polymerase but dependent on sub-sampling depth (sub-sampled to 1000 =
0.51 (0.4 - 0.79) (median (25\% - 75\% quantile)), sub-sampled to 5000 =
0.43 (0.33 - 0.63), sub-sampled to 10000 = 0.4 (0.24 - 0.43)) {[}Figure
3A{]}. Second, for the mocks, where data is available, there were larger
difference between 20x and 25x (sub-sampled to 1000 = 0.88 (0.42 -
0.91)) {[}Figure 3B{]}. However, these differences between the next 5
cycle increment do not persist once 25 cycles are reached {[}Figure
3{]}.

Using PERMANOVA to test for differences within HiFi DNA polymerase
groups based on cycle number, only Phusion had cycle dependent
differences at 1000 and 5000 sub-sampling depth (P-value = 0.03 and
0.01, respectively). Phusion was one of only two HiFi DNA polymerases
that managed to have samples for the 1000 sub-sampling depth at 15
cycles. Next, we assessed whether there were any major differences
between 5 cycle increments within each sample. We found that there was
no detectable difference in Bray-Curtis index when comparing to the
previous 5 cycle increment (P-value \textgreater{} 0.05). However,
Phusion at 1000 sub-sampling depth had a P-value = 0.02 before multiple
comparison correction. It should be noted that at higher sub-sampling
depths these differences in Bray-Curtis indices disappear for both
differences in cycle number and within 5 cycle increments within an
individual.

\textbf{\emph{Sequence Error is Dependent on both HiFi DNA Polymerase
and Cycle Number:}} Differences by HiFi DNA polymerase in the median
average per base error varied without a clear pattern across
sub-sampling depth {[}Table S5{]}. Generally, the highest per base
median average error rates were for the Kappa HiFI DNA polymerase
{[}Figure 4{]}. This error rate was minimally affected by both the
pre-cluster step and chimera removal by VSEARCH {[}Figure 4{]}. There
were small differences in the per base error rate between the various
HiFi DNA polymerase at lower cycle numbers and larger differences at
higher cycle number with Platinum having the largest differences of all
the HiFi DNA polymerase {[}Figure 4B-C and Table S6{]}.

The total sequences with at least one error had multiple differences at
different cycle numbers and was mostly alleviated by the use of the
pre.cluster step {[}Figure S1{]}. Major differences before this
pre.cluster step were driven by large differences in
Accuprime\textsuperscript{TM} and Platinum versus the other HiFi DNA
polymerase tested {[}Figure S1 \& Table S7 \& S8{]}. Although
Accuprime\textsuperscript{TM} had the lowest per base error rate it had
the largest number of sequences with at least one error, regardless of
cycle number or sub-sampling depth {[}Figure S1{]}. However, this
increased number of sequences with an error can be drastically lowered
with existing bioinformatic approaches {[}Figure S1{]}. Investigation of
whether there were HiFi DNA polymerase dependent effects on base
substitution found that there were generally no biases in the types of
substitution made {[}Figure S2{]}.

\textbf{\emph{Chimeric Sequences Correlate with OTUs and are HiFi DNA
Polymerase Dependent:}} After chimera removal using VSEARCH and removal
of sequences that did not classify as bacteria we assessed the
percentage of sequences that were still chimeric within our mock
community. At all levels of sub-sampling and cycle number there were
significant differences between the HiFi DNA polymerase used (P-value
\textless{} 0.05) {[}Table S9{]}. Differences between Platinum and all
other HiFi DNA polymerases accounted for the vast majority of these
differences independent of cycle number and sub-sampling depth when
using a Dunn's post-hoc test {[}Table S10{]}. Generally, across
sub-sampling depth and cycle number Accuprime\textsuperscript{TM} had
the lowest chimera prevalence of all the HiFi DNA Polymerases regardless
of whether pre.cluster or VSEARCH had been used {[}Figure 5{]}.

For all DNA polymerase, a positive correlation was observed between
chimeric sequences and number of OTUs, with this correlation being
strongest for Accuprime, Platinum and Phusion HiFi DNA Polymerase
{[}Figure 6{]}. In general, the R\textsuperscript{2} value between the
number of OTUs and chimeric sequences did not change from the use of
pre.cluster and VSEARCH {[}Figure 6{]}. Taken together, this data
suggests that a strong correlation exists between the number of OTUs and
the prevalence of chimeric sequences.

>>>>

\textbf{\emph{The Number of OTUs is Dependent on the HiFi DNA Polymerase
Used:}} In order to compare the number of OTUs across individuals a
Z-score normalization by individual, by cycle number, was employed on
the number of OTUs data. After normalization, we identified that there
was a HiFi DNA polymerase dependent difference that was consistent
across sub-sampling {[}Figure 1{]}. Lower cycle numbers (15-20) tended
to result in less differences between HiFI DNA polymerase while cycle
numbers of 25, 30, and 35 tended to have more distinct differences
{[}Figure 1{]}. All sub-sampling levels were significantly different
only at 35 cycles (P-value \textless{} 0.0001) {[}Table S1{]}. There
were differences in HiFi DNA polymerase at 25 and 30 cycles but the
sub-sampling depth had to be 5000 or higher (P-value \textless{} 0.05)
{[}Table S1{]}. Platinum HiFi DNA polymerase was the main driver of the
differences observed across all sub-sampling depths at 35 cycles, based
on a Tukey post-hoc test (P-value \textless{} 0.05) {[}Table S2{]}.

@@

\textbf{\emph{Minimal Bray-Curtis Differences are Detected by Cycle
Number:}} Overall, there was little difference between each respective
5-cycle increment at higher cycle numbers, for both fecal and mock
samples, and this was consistent across the different sub-samplings
{[}Figure 3{]}. There were only two differences between 5-cycle
increments that were identified. First, there were differences for the
same fecal sample between 20 versus 25 cycles that was independent of
HiFi DNA polymerase but dependent on sub-sampling depth (sub-sampled to
1000 = 0.51 (0.4 - 0.79) (median (25\% - 75\% quantile)), sub-sampled to
5000 = 0.43 (0.33 - 0.63), sub-sampled to 10000 = 0.4 (0.24 - 0.43))
{[}Figure 3A{]}. Second, for the mocks, where data is available, there
were larger difference between 20 and 25 cycles (sub-sampled to 1000 =
0.88 (0.42 - 0.91)) {[}Figure 3B{]}. However, these differences between
the next 5-cycle increment do not persist once 25 cycles are reached
{[}Figure 3{]}.

Using PERMANOVA to test for differences within HiFi DNA polymerase
groups based on cycle number, only Phusion had cycle dependent
differences at 1000 and 5000 sub-sampling depth (P-value = 0.03 and
0.01, respectively). Phusion was one of only two HiFi DNA polymerases
that managed to have samples for the 1000 sub-sampling depth at 15
cycles. Next, we assessed whether there were any major differences
between 5-cycle increments within each sample. We found that there was
no detectable difference in Bray-Curtis index when comparing to the
previous 5-cycle increment (P-value \textgreater{} 0.05). However,
Phusion at 1000 sub-sampling depth had a P-value = 0.02 before multiple
comparison correction. It should be noted that at higher sub-sampling
depths these differences in Bray-Curtis indices disappear for both
differences in cycle number and within 5-cycle increments within an
individual.

\textbf{\emph{Sequence Error is Dependent on both HiFi DNA Polymerase
and Cycle Number:}} Differences by HiFi DNA polymerase in the median
average per base error varied without a clear pattern across
sub-sampling depth {[}Table S5{]}. Generally, the highest per base
median average error rates were for the Kappa HiFi DNA polymerase
{[}Figure 4{]}. This error rate was minimally affected by both the
`pre.cluster' step and chimera removal by VSEARCH {[}Figure 4{]}. There
were small differences in the per base error rate between the various
HiFi DNA polymerase at lower cycle numbers and larger differences at
higher cycle number with Platinum having the largest differences of all
the HiFi DNA polymerase {[}Figure 4B-C and Table S6{]}.

The total sequences with at least one error had multiple differences at
different cycle numbers and was mostly alleviated by the use of the
`pre.cluster' step {[}Figure S1{]}. Major differences before this
`pre.cluster' step were driven by large differences in
Accuprime\textsuperscript{TM} and Platinum versus the other HiFi DNA
polymerase tested {[}Figure S1 \& Table S7 \& S8{]}. Although
Accuprime\textsuperscript{TM} had the lowest per base error rate it had
the largest number of sequences with at least one error, regardless of
cycle number or sub-sampling depth {[}Figure S1{]}. However, this
increased number of sequences with an error can be drastically lowered
with existing bioinformatic approaches {[}Figure S1{]}. Investigation of
whether there were HiFi DNA polymerase dependent effects on base
substitution found that there were generally no biases in the types of
substitution made {[}Figure S2{]}.

\textbf{\emph{Chimeric Sequences are HiFi DNA Polymerase Dependent and
Correlate with Number of OTUs:}} After chimera removal using VSEARCH and
removal of sequences that did not classify as bacteria, we assessed the
percentage of sequences that were still chimeric within our mock
community. At all levels of sub-sampling and cycle number there were
significant differences between the HiFi DNA polymerase used (P-value
\textless{} 0.05) {[}Table S9{]}. Differences between Platinum and all
other HiFi DNA polymerases accounted for the vast majority of these
differences independent of cycle number and sub-sampling depth when
using a Dunn's post-hoc test {[}Table S10{]}. Generally, across
sub-sampling depth and cycle number Accuprime\textsuperscript{TM} had
the lowest chimera prevalence of all the HiFi DNA Polymerases regardless
of whether `pre.cluster' or chimera removal with VSEARCH had been used
{[}Figure 5{]}.

For all DNA polymerase, a positive correlation was observed between
chimeric sequences and number of OTUs, with this correlation being
strongest for Accuprime, Platinum and Phusion HiFi DNA Polymerase
{[}Figure 6{]}. The R\textsuperscript{2} value between the number of
OTUs and chimeric sequences did not change from the use of `pre.cluster'
and the removal of chimeras with VSEARCH {[}Figure 6{]}. Taken together,
this data suggests that a strong correlation exists between the number
of OTUs and the prevalence of chimeric sequences.
